---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## addindicators

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/addindicators/branch/master/graph/badge.svg?token=RlTJbum32D)](https://codecov.io/gh/impact-initiatives/addindicators)


  <!-- badges: end -->



## Overview
`addindicators` is designed for creating and checking (occasionally) composite indicators such as Food Consumption Indicators and others.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/addindicators")
```


## Examples

```{r include=T,eval=T}
library(addindicators)
df <- addindicators_MSNA_template_data
```

### Example:: Add Food Consumption Score (FCS)
``` {r eval=T,warning=F}
df_with_fcs <- df %>% add_fcs(
  cutoffs = "normal 21.5-35",
  fcs_cereal = "fs_fcs_cereals_grains_roots_tubers",
  fcs_legumes = "fs_fcs_beans_nuts",
  fcs_veg = "fs_fcs_vegetables_leaves",
  fcs_fruit = "fs_fcs_fruit",
  fcs_meat = "fs_fcs_meat_fish_eggs",
  fcs_dairy = "fs_fcs_dairy",
  fcs_sugar = "fs_fcs_sugar",
  fcs_oil = "fs_fcs_oil_fat_butter"
 )
df_with_fcs %>% dplyr::select(uuid,fcs_score,fcs_cat,fcs_weight_cereal1,fcs_weight_legume2,
                       fcs_weight_dairy3,fcs_weight_meat4,fcs_weight_veg5,
                       fcs_weight_fruit6,fcs_weight_oil7,fcs_weight_sugar8) %>% head(20)
```

### Example:: Add Household Hunger Scale (HHS)
``` {r, eval=TRUE, warning=F}
df_with_hhs <- df_with_fcs %>% add_hhs(
     hhs_nofoodhh_1 = "fs_hhs_nofood_yn",
     hhs_nofoodhh_1a = "fs_hhs_nofood_freq",
     hhs_sleephungry_2 = "fs_hhs_sleephungry_yn",
     hhs_sleephungry_2a = "fs_hhs_sleephungry_freq",
     hhs_alldaynight_3 = "fs_hhs_daynoteating_yn",
     hhs_alldaynight_3a = "fs_hhs_daynoteating_freq",
     yes_answer = "yes",
     no_answer = "no",
     rarely_answer = "rarely_1_2",
     sometimes_answer = "sometimes_3_10",
     often_answer = "often_10_times"
)
df_with_hhs %>% dplyr::select(uuid, hhs_comp1, hhs_comp2,hhs_comp3,
                       hhs_score, hhs_cat_ipc, hhs_cat, hh_size) %>% head(20)
```

### Example:: Add Livelihood Coping Strategy score (LCSI)
``` {r, eval=TRUE, warning=F}
df_with_lcsi <- df_with_hhs %>% add_lcsi(
 lcsi_stress_vars = c("liv_stress_lcsi_1", "liv_stress_lcsi_2", "liv_stress_lcsi_3", "liv_stress_lcsi_4"),
 lcsi_crisis_vars = c("liv_crisis_lcsi_1", "liv_crisis_lcsi_2", "liv_crisis_lcsi_3"),
 lcsi_emergency_vars = c("liv_emerg_lcsi_1", "liv_emerg_lcsi_2", "liv_emerg_lcsi_3"),
 yes_val = "yes",
 no_val = "no_had_no_need",
 exhausted_val = "no_exhausted",
 not_applicable_val = "not_applicable"
)
df_with_lcsi %>% dplyr::select(uuid, lcsi_cat,lcsi_cat_exhaust,lcsi_cat_yes)%>% head(20)
```

### Example:: Add Reduced Household Coping Strategy score (rCSI)
``` {r, eval=TRUE, warning=F}
df_with_rcsi <- df_with_lcsi %>% add_rcsi(
  rCSILessQlty = "rCSILessQlty",
  rCSIBorrow = "rCSIBorrow",
  rCSIMealSize = "rCSIMealSize",
  rCSIMealAdult = "rCSIMealAdult",
  rCSIMealNb = "rCSIMealNb",
  new_colname = "rcsi"
)
df_with_rcsi %>% dplyr::select(uuid,rcsi_score,rcsi_cat)%>% head(20)
```

### Example:: Add Food Consumption Matrix (FCM)
**Notice that these functions are also pipable**
``` {r, eval=TRUE, warning=F}
df_with_fcm <- df_with_rcsi %>%
  add_fcm_phase(
   fcs_column_name = "fcs_cat",
   rcsi_column_name = "rcsi_cat",
   hhs_column_name = "hhs_cat_ipc",
   fcs_categories_acceptable = "Acceptable",
   fcs_categories_poor = "Poor",
   fcs_categories_borderline = "Borderline",
   rcsi_categories_low = "No to Low",
   rcsi_categories_medium = "Medium",
   rcsi_categories_high = "High",
   hhs_categories_none = "None",
   hhs_categories_little = "Little",
   hhs_categories_moderate = "Moderate",
   hhs_categories_severe = "Severe",
   hhs_categories_very_severe = "Very Severe"
)
df_with_fcm %>% dplyr::select(uuid,fc_cell,fc_phase)%>% head(20)
```

### Example:: Add FEWSNET Food Consumption-Livelihood Matrix (FCLCM)
**Notice that these functions are also pipable**
``` {r, eval=TRUE, warning=F}
df_with_fclcm <- df_with_fcm %>% ## Taken from previous Example
  add_fclcm_phase()
df_with_fclcm %>% dplyr::select(uuid,fclcm_phase)%>% head(20)
```
