---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## addindicators

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/addindicators/branch/master/graph/badge.svg?token=RlTJbum32D)](https://codecov.io/gh/impact-initiatives/addindicators)


  <!-- badges: end -->



## Overview
`addindicators` is designed for creating and checking (occasionally) composite indicators like Food Consumption Score (FCS) or Household Hunger Score (HHS) 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/addindicators")
```


## example

The recommendated steps for calculating FCS and HHS are given below. 

**Step 1:** Reading the dataset
``` {r, eval=FALSE}
df <- read.csv("[file path].csv")
```

**Step 2:** Adding HHS

``` {r, eval=FALSE}
df_with_hhs <- df %>% add_hhs(
  new_colname = "hhs",
  hhs1a = "no_food_to_eat_because_of_lack_of_resources",
  hhs1b = "how_often_no_food_to_eat_because_of_lack_of_resources",
  hhs2a = "hh_memeber_sleep_hungry",
  hhs2b = "how_often_hh_memeber_sleep_hungry",
  hhs3a = "hh_member_go_a_whole_day_and_night_without_eating",
  hhs3b = "how_often_hh_member_go_a_whole_day_and_night_without_eating",
  frequency_choice = c("rarely_1_2_times", "sometimes_3_10_times", "often_10_plus_times")
)
```

**Step 2:** Adding FCS

``` {r, eval=FALSE}
df_with_hhs_and_fcs <- df_with_hhs %>% make_fcs(
  var_name = "fcs",
  cereals = "cereals_grains_roots_tubers",
  pulses = "beans_legumes_pulses_nuts",
  dairy = "milk_dairy_products",
  meat = "meat_fish_eggs",
  vegetables = "vegetables",
  fruits = "fruite",
  oil = "oil_fat_butter",
  sugar = "sugar_sugary_food"
)
```




