---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## addindicators

<!-- badges: start -->
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)
[![R-CMD-check](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/addindicators/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/addindicators/branch/master/graph/badge.svg?token=RlTJbum32D)](https://codecov.io/gh/impact-initiatives/addindicators)
<!-- badges: end -->



## Overview
`addindicators` is designed for creating and checking (occasionally) composite indicators such as Food Consumption Indicators and others.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/addindicators")
```


## Examples

```{r include=T,eval=T}
library(addindicators)
df <- addindicators_MSNA_template_data
```

### Example:: Add Food Consumption Score (FCS)
``` {r eval=T,warning=F}
df_with_fcs <- df %>% add_fcs(
  cutoffs = "normal",
  fsl_fcs_cereal = "fs_fcs_cereals_grains_roots_tubers",
  fsl_fcs_legumes = "fs_fcs_beans_nuts",
  fsl_fcs_veg = "fs_fcs_vegetables_leaves",
  fsl_fcs_fruit = "fs_fcs_fruit",
  fsl_fcs_meat = "fs_fcs_meat_fish_eggs",
  fsl_fcs_dairy = "fs_fcs_dairy",
  fsl_fcs_sugar = "fs_fcs_sugar",
  fsl_fcs_oil = "fs_fcs_oil_fat_butter"
)
df_with_fcs %>%
  dplyr::select(
    uuid,  fsl_fcs_score, fsl_fcs_cat, fcs_weight_cereal1, fcs_weight_legume2,
    fcs_weight_dairy3, fcs_weight_meat4, fcs_weight_veg5,
    fcs_weight_fruit6, fcs_weight_oil7, fcs_weight_sugar8
  ) %>%
  head(20)
```

### Example:: Add Household Hunger Scale (HHS)
``` {r, eval=TRUE, warning=F}
df_with_hhs <- df_with_fcs %>% add_hhs(
  hhs_nofoodhh_1 = "fs_hhs_nofood_yn",
  hhs_nofoodhh_1a = "fs_hhs_nofood_freq",
  hhs_sleephungry_2 = "fs_hhs_sleephungry_yn",
  hhs_sleephungry_2a = "fs_hhs_sleephungry_freq",
  hhs_alldaynight_3 = "fs_hhs_daynoteating_yn",
  hhs_alldaynight_3a = "fs_hhs_daynoteating_freq",
  yes_answer = "yes",
  no_answer = "no",
  rarely_answer = "rarely_1_2",
  sometimes_answer = "sometimes_3_10",
  often_answer = "often_10_times"
)
df_with_hhs %>%
  dplyr::select(
    uuid, hhs_comp1, hhs_comp2, hhs_comp3,
    hhs_score, hhs_cat_ipc, hhs_cat, hh_size
  ) %>%
  head(20)
```

### Example:: Add Livelihood Coping Strategy score (LCSI)
``` {r, eval=TRUE, warning=F}
df_with_lcsi <- df_with_hhs %>% add_lcsi(
  lcsi_stress_vars = c("liv_stress_lcsi_1", "liv_stress_lcsi_2", "liv_stress_lcsi_3", "liv_stress_lcsi_4"),
  lcsi_crisis_vars = c("liv_crisis_lcsi_1", "liv_crisis_lcsi_2", "liv_crisis_lcsi_3"),
  lcsi_emergency_vars = c("liv_emerg_lcsi_1", "liv_emerg_lcsi_2", "liv_emerg_lcsi_3"),
  yes_val = "yes",
  no_val = "no_had_no_need",
  exhausted_val = "no_exhausted",
  not_applicable_val = "not_applicable"
)
df_with_lcsi %>%
  dplyr::select(uuid, lcsi_cat, lcsi_cat_exhaust, lcsi_cat_yes) %>%
  head(20)
```

### Example:: Add Reduced Household Coping Strategy score (rCSI)
``` {r, eval=TRUE, warning=F}
df_with_rcsi <- df_with_lcsi %>% add_rcsi(
  rCSILessQlty = "rCSILessQlty",
  rCSIBorrow = "rCSIBorrow",
  rCSIMealSize = "rCSIMealSize",
  rCSIMealAdult = "rCSIMealAdult",
  rCSIMealNb = "rCSIMealNb",
  new_colname = "rcsi"
)
df_with_rcsi %>%
  dplyr::select(uuid, rcsi_score, rcsi_cat) %>%
  head(20)
```

### Example:: Add Food Consumption Matrix (FCM)
**Notice that these functions are also pipable**
``` {r, eval=TRUE, warning=F}
df_with_fcm <- df_with_rcsi %>%
  add_fcm_phase(
    fcs_column_name = "fsl_fcs_cat",
    rcsi_column_name = "rcsi_cat",
    hhs_column_name = "hhs_cat_ipc",
    fcs_categories_acceptable = "Acceptable",
    fcs_categories_poor = "Poor",
    fcs_categories_borderline = "Borderline",
    rcsi_categories_low = "No to Low",
    rcsi_categories_medium = "Medium",
    rcsi_categories_high = "High",
    hhs_categories_none = "None",
    hhs_categories_little = "Little",
    hhs_categories_moderate = "Moderate",
    hhs_categories_severe = "Severe",
    hhs_categories_very_severe = "Very Severe"
  )
df_with_fcm %>%
  dplyr::select(uuid, fc_cell, fc_phase) %>%
  head(20)
```

### Example:: Add FEWSNET Food Consumption-Livelihood Matrix (FCLCM)
**Notice that these functions are also pipable**
``` {r, eval=TRUE, warning=F}
df_with_fclcm <- df_with_fcm %>% ## Taken from previous Example
  add_fclcm_phase()
df_with_fclcm %>%
  dplyr::select(uuid, fclcm_phase) %>%
  head(20)
```


### Example:: Review of indicators

The logic behind *review_variables* is to compare the results from 2 codes to create the composite variable. 

In this example, the Food Consumption Score from the first example will be compared.

```{r}
review_df <- addindicators_MSNA_template_data %>% add_fcs(
  cutoffs = "normal",
  fsl_fcs_cereal = "fs_fcs_cereals_grains_roots_tubers",
  fsl_fcs_legumes = "fs_fcs_beans_nuts",
  fsl_fcs_veg = "fs_fcs_vegetables_leaves",
  fsl_fcs_fruit = "fs_fcs_fruit",
  fsl_fcs_meat = "fs_fcs_meat_fish_eggs",
  fsl_fcs_dairy = "fs_fcs_dairy",
  fsl_fcs_sugar = "fs_fcs_sugar",
  fsl_fcs_oil = "fs_fcs_oil_fat_butter"
)
```

The new results and the results to be reviewed are bound together by the *uuid*.

```{r}
binded_df <- df_with_fcs %>%
  dplyr::full_join(review_df, by = "uuid")
```

There are 2 functions to review:
  - review_one_variable, to review only one variable
  - review_variables, a wrapper around review_one_variable to be able to review several variables.

#### review_one_variable
```{r}
review_one_variable <- review_one_variable(binded_df,
  column_to_review = "fsl_fcs_cat.x",
  column_to_compare_with = "fsl_fcs_cat.y"
)

review_one_variable$review_check_fsl_fcs_cat.x %>% mean()
```


```{r}
review_one_variable$review_comment_fsl_fcs_cat.x %>% table(useNA = "ifany")
```

#### review_variables
```{r}
review_results <- review_variables(binded_df,
  columns_to_review = c("fsl_fcs_score.x", "fsl_fcs_cat.x"),
  columns_to_compare_with = c("fsl_fcs_score.y", "fsl_fcs_cat.y")
)

review_results$review_table %>%
  dplyr::group_by(variable) %>%
  dplyr::summarise(prop_correction = mean(review_check))

review_results$review_table %>%
  dplyr::group_by(variable, review_comment) %>%
  dplyr::tally(sort = T)
```

#### Examples when differences exists
```{r}
test_categorical <- data.frame(
  test = c(
    "test equality",
    "test difference",
    "test Missing in y",
    "test Missing in x",
    "test equality missing in both"
  ),
  var_x = c("A", "B", "C", NA, NA),
  var_y = c("A", "A", NA, "D", NA),
  uuid = letters[1:5]
)
review_one_variable(test_categorical,
  column_to_review = "var_x",
  column_to_compare_with = "var_y"
)
```

## Code of Conduct

Please note that the addindicators project is released with a [Contributor Code of Conduct](https://impact-initiatives.github.io/addindicators/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
